\documentclass[12pt]{letter}
\usepackage[a4paper,left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=black, filecolor=blue, citecolor=black, urlcolor=blue}

\begin{document}
\signature{V. V. Bashun \\ Assistant Professor \\ HSE University \\ \url{vbashun@hse.ru}}

\begin{letter}{}
\opening{Dear Admission Committee,}

It is my pleasure to recommend Aleksei Shadrunov for admission to the master's program at your university. I have had the opportunity to work closely with Aleksei as a lecturer for several courses during his bachelor's degree, and as an academic supervisor for his thesis. I can attest to his persistence and professional qualities.

Aleksei has been a diligent and bright student since he attended the Project Seminar on Information Security in his first year. Motivated by his thirst for knowledge, he quickly grasped the concepts of web application security, vulnerabilities and attacks. In his second year, he deepened his understanding of these topics and proved himself to be a skilled PHP developer, not only understanding the syntax but also adhering to secure software development practices.

In his fourth year of study, Aleksei developed a keen interest in security of software development and operations. This is the area he focused on in his bachelor's thesis project which explored the topic of hardening of containerized applications. Through this project, Aleksei demonstrated his ability to work with scientific literature and conduct practical research.

Apart from succeeding academically, Aleksei also played a significant role in the life of his faculty and university. As the head of a group, he was responsible for organising and communicating within and outside of classes, always available to provide assistance to other students and professors. Moreover, Aleksei was an active member of several student volunteer organisations, which helped him to master his interpersonal skills and grow outside the professional field.

Your program includes various aspects of information security, and I believe that Aleksei would be able to succeed in it and achieve his professional and academic goals. I encourage you to consider his application favourably.

\closing{Sincerely,}


\end{letter}
\end{document}

