\documentclass[12pt]{letter}
\usepackage[a4paper,left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=black, filecolor=blue, citecolor=black, urlcolor=blue}

\begin{document}
\signature{Yaroslav E. Meshcheryakov \\ Associate Professor at the Department of Cyber-Physical Systems Information Security \\ HSE University \\ \url{ymeshcheryakov@hse.ru}}

\begin{letter}{}
\opening{Dear Admission Committee,}

I am pleased to write this letter of recommendation for Aleksei Shadrunov for admission to the master's program at your university. I have known him since September 2021 and can highly recommend him as both an excellent student and a qualified professional.

Aleksei was initially introduced to me during his participation in the “Digitalskills” Industrial Championship. Under my supervision, he demonstrated outstanding ability to focus on challenging tasks, mobilise his knowledge and apply hands-on experience within the time constraints of the competition. Aleksei's diligence and resilience in his preparation allowed him to achieve a prize-winning position for his university.

Later in his degree, Aleksei attended my courses on System Programming and Electronics and Electric Engineering. He was an attentive and active participant in the class, demonstrating interest in the architecture of modern computing systems and low-level development. In addition to completing his assignments successfully, Aleksei regularly presented on topics that he was passionate about and participated in group work in the laboratory.

I have been pleased to observe Aleksei's passion for exploration during our classes over these past few years, and I am confident in his ability to succeed in your master's program. I would appreciate it if you could consider his application favourably.

\closing{Sincerely,}


\end{letter}
\end{document}

