\newpage
\section*{Acknowledgments}
% \addcontentsline{toc}{section}{Annotation}

I would like to express my gratitude to my supervisor, V.V. Bashun, for his assistance in navigating the field of security research, helping to formulate the topic of this study, providing guidance on relevant materials. He has encouraged me to dedicate more time and effort to this research, inspiring me to strive for greater goals. Not only has he provided valuable advice on this paper, but he has taught me the majority of the relevant material throughout my bachelor's degree, igniting my passion for container technologies. His encouragement and support have been pivotal in shaping my academic journey.

Next, I am grateful to my dear friends from university, Alisa Kulishenko and Andrei Krechetov, for making this journey bearable and celebrating the highs and lows along the way. Special thanks to Maksim Emelyanenko for his assistance with the \LaTeX \ formatting. I am also deeply grateful to Alena Baranova for her invaluable advice on working with texts, the English language, and for sharing her knowledge of academic traditions. But, more importantly, she has been a constant source of support for me throughout the years.

I am grateful to my former colleagues at Kaspersky Lab for providing me with the opportunity to learn about best practices in working with containers and managing vulnerabilities in industry. Finally, I want to thank my family for their support and strong belief in me.


\pagebreak